---
title: "tumex5: Five TCGA tumors' expression data for workflow experimentation"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup,echo=FALSE}
suppressPackageStartupMessages({
suppressMessages({
library(rhdf5client)
library(tumex5)
library(BiocOncoTK)
library(restfulSE)
})
})
```

# Introduction

We would like to demonstrate how workflows can employ
HSDS and (RESTful) SummarizedExperiment instances.

We retrieved five tumors' expression data using
curatedTCGAData.  These were converted to HDF5 SummarizedExperiment
instances, and the HDF5 representations of quantifications
were ported to HSDS using hsload.

The following function produces a 'long' dataset for a
given tumor type, binding the MSIsensor scores of Ding et al 2018.

```{r def}
 exprByMSI = function(tumcode, genesym, alias) {
  if (missing(alias)) alias=genesym
  ob = paste0(tumcode, "rnagn")
  ex = get(ob)
  ex = BiocOncoTK::bindMSI(ex)
  data.frame(
   patient_barcode=colnames(ex),
   acronym=tumcode,
   symbol = genesym,
   alias = alias,
   log2ex=log2(as.numeric(SummarizedExperiment::assay(ex[genesym,]))+1),
   msicode = ifelse(ex$msiTest >= 4, ">=4", "<4"))
 }
head(exprByMSI("COAD", "CD8A"))
```

# Application

We define a small set of genes indicative of immune
infiltration of tumor.  We then build a data.frame
instance with tumor codes and gene symbols constituting
all combinations of tumor type and gene.
```{r opts}
infilGenes = c(`PD-L1`="CD274", `PD-L2`="PDCD1LG2", CD8A="CD8A")
tumcodes = c("COAD", "STAD", "UCEC")
combs = expand.grid(tumcode=tumcodes, ali=names(infilGenes),
    stringsAsFactors=FALSE)
combs$sym = infilGenes[combs$ali]
```

The job is completed by building a long data.frame
for all combinations and visualizing.

```{r lktab, fig.height=5}
 allshow = lapply(1:nrow(combs), function(x) exprByMSI(combs$tumcode[x],
    combs$sym[x], combs$ali[x]))
 
 rr = do.call(rbind, allshow)
 
 library(ggplot2)
 ggplot(rr,
    aes(msicode, log2ex)) + geom_boxplot() +
    facet_grid(acronym~alias) +
    ylab("log2(normalized expr. + 1)") +
    xlab("microsatellite instability score")
```
